# Project Settings

```{r}
#| label: zam_analysis_file_creation
#| message: false
#| echo: fenced
#| warning: false
require(Hmisc)
require(qreport)
require(qs)
require(here)
require(ggplot2)
require(wesanderson)
require(cowplot)
require(lubridate)
require(gt)
source(here("R/paths.r"))
source(here("R/plot.r"))
options(prType = "html")
hookaddcap()
```

## Population

```{r}
#| label: zam_dh_vl_pop
#| warning: false
#| message: false
#| column: page
#| tbl-cap: Overview of the village population data
#| tbl-width: 100%
#| tbl-align: "center"
#| tbl-font-size: 12 
zam_dh_vl_pop <- qread(here(PATH_ZAM_DH_VL_POP))

zam_dh_vl_pop |>
    gt() |>
    tab_header(
        title = "Zambia DH VL Population",
        subtitle = "Overview of the village population data"
    ) |>
    fmt_number(
        columns = where(is.numeric),
        decimals = 0
    ) |>
    cols_align(
        align = "center",
        columns = everything()
    ) |>
    tab_style(
        style = list(
            cell_fill(color = "lightblue"),
            cell_text(weight = "bold")
        ),
        locations = cells_column_labels()
    ) |>
    opt_row_striping() |>
    opt_table_font(
        font = list(google_font("Roboto"))
    ) |>
    tab_options(
        table.width = pct(100),
        table.border.top.color = "black",
        table.border.bottom.color = "black",
        column_labels.border.bottom.color = "black",
        column_labels.border.bottom.width = px(3)
    )

```


::: {.callout-warning}
## TODO
The village list for each HFCA is still pending. We are waiting for Zambia to send this information back.
:::


## Date of project

The project started in 2022 and ended in 2023.

The DHIS2 data were collected from 2022 to 2023.

The survey data were collected from 2022 to 2023.

## Intervention group and control group settgings

Two HFCA were selected based on the protocal of the 1,7 mCTRC project.

In this analysis, we will focus on the following HFCA:

-   Kankumba
-   Lukwipa


::: {.callout-note}
## Intervention and Control Group Definition

**TO BE CONFIRMED BY ZAMBIA**

The selection process for the intervention and control groups was as follows:

1. Two Health Facility Catchment Areas (HFCAs) were selected based on the protocol of the 1,7 mCTRC project.
2. These HFCAs were then randomly assigned to either the intervention or control group.

The final group assignments are:

- **Intervention Group**: Kankumba HFCA
- **Control Group**: Lukwipa HFCA

The intervention group received the 1,7-mRCT (monthly reactive case treatment) intervention, while the control group continued with standard malaria control measures.

This clear definition of intervention and control groups allows for a robust comparison of the effectiveness of the 1,7-mRCT intervention in reducing malaria transmission.
:::

## Geographical distribution
